---
title: "Analyzing Transmission impact on MPG"
author: "Raj Kumar"
date: "10/29/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
##Loading Important Libraries and datasets

#Loading Libraries
library(randomForest)
library(datasets)
library(car)
library(caret)

#loading dataset
car_data <- mtcars

```

## Preparing data

*Checking if NAs value are present in the dataset  
*Checking if any outliers exist and if it exist then impute it with mean values.  


```{r, echo=TRUE}
#looking at summary of the variables
summary(car_data)
```
We can see that is no NAs in the dataset

```{r, echo=TRUE}
##Checking distribution of MPG for each transmission type
boxplot(car_data$mpg~car_data$am, ylab = "Miles/Gallon", xlab = "Transmission Type")

#Checking if outliers exists
par(mfrow = c(2,2))
boxplot(car_data$mpg)
boxplot(car_data$disp)
boxplot(car_data$drat)
boxplot(car_data$qsec) 
par(mfrow = c(1,2))
boxplot(car_data$wt)  
boxplot(car_data$hp)

```

Found some outliers in qsec,wt and hp  
Imputing outliers with mean value

```{r, echo=TRUE}
car_data[which(car_data$hp >(quantile(mtcars$hp,0.75) + 1.5*(IQR(mtcars$hp)))),c("hp")] <- mean(car_data$hp)
car_data[which(car_data$qsec >(quantile(mtcars$qsec,0.75) + 1.5*(IQR(mtcars$qsec)))),c("qsec")] <- mean(car_data$qsec)
car_data[which(car_data$wt >(quantile(mtcars$wt,0.75) + 1.5*(IQR(mtcars$wt)))),c("wt")] <- mean(car_data$wt)
```

## Fitting Model

Checking if Transmission Affect MPG

```{r , echo=TRUE}
##Fitting a regression model
model <- lm(mpg ~ am , data = car_data)
summary(model)
confint(model)
```

Looking at the summary we can say that difference is significant with confidence of 95%  as p value is 0.000285 (p<0.001).  
Also R-Square value is 0.36 which means that only 36 % of the model is explained by transmission.  
That means there are other variables which is also impacting the MPG of the car  

```{r , echo=TRUE}
 
##Checking Impact of other variables on the MPG
model_mul <- lm(mpg ~ . , data = car_data)
summary(model_mul)

##Checking for multicollinearity
vif(model_mul)
##Removing Cyl as it has a high variable inflation factor

##Using Step function to get the most important variable
model_step<-step(model_mul,direction = "both")
```

Including am and the variables suggested by step function

```{r , echo=TRUE}
model_final <- lm(mpg ~ disp + wt + carb + am, data = car_data)
summary(model_final)
```

Now we can see that  model can explain about 85 % of the variance due to 4 variables.

## Model Diagnostics
Comparing the two models and checking the Residual plots

```{r , echo=TRUE}
##Comparing the two models
anova(model,model_final)


#####looking at residual plots
par(mfrow = c(2,2))
plot(model_final)
```

the plot looks good.  
residuals are randomly distributed 
